### 2021-06-14 HW-Gravity/gravity1/gravity1.tex
Add minor tweaks to various notes
Finish gravity1 homework, add mathematica script

% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

@@

@@

	From now on we shall focus on the $p^2 \ne 0$ situation. Note that we can complete the square on the right hand side of the $k^\mu$ conservation such that:
	\begin{equation}
		k_\mu + p_\mu a^2
		= p_\mu \pqty{
			z^2 + (x - a)^2
		},
	\quad \Delta = p_\mu a^\mu
	\end{equation}
	For $p^\mu \ne 0$, we can always find some $c_\mu$ such that $c_\mu p^\mu \ne 0$, therefore we have:
	\begin{equation}
		z^2 + \pqty{x - a}^2
		= \frac{c_\mu k^\mu}{c_\mu p^\mu}
			+ a^2
		= \pqty{\frac{L}{2}}^{\!\!2}
		= \mrm{const.},
	\quad \Delta = p_\mu a^\mu
	\end{equation}

@@

	This confirms our observation above, and further reveals that $\gamma$ lies in a ``plane'' in the $\mbb{R}^d$ subspace consisting of the $x^\mu$ coordinates. The behavior of $\gamma$ is sensitive to the sign of $\delta$ and $p^2$; more specifically,
	\begin{enumerate}
	\item $\delta = -1$, i.e.~for timelike $\gamma$, we must have $p^2 < 0$, and $z \ge \frac{1}{\abs{\norm{p}}} = \frac{L}{2} > 0$, namely $z$ is bounded from below. This means that timelike geodesic can never reach the asymptotic boundary $z \to 0$. In this case, $\gamma$ is a section of the one-sheet hyperboloid. 
	\item $\delta = +1$, i.e.~for spacelike $\gamma$, we can have $p^2 > 0$ or $p^2 < 0$.
		\begin{enumerate}
		\item For $p^2 > 0$, again we have $\frac{L}{2} = \frac{1}{\norm{p}} > 0$, and $\gamma$ is again a cross section of the one-sheet hyperboloid. However, now we have $z \le \frac{L}{2}$, namely $z$ is bounded from above, and:
		\begin{equation}
			z\to 0,
		\quad
			x^\mu
			= a^\mu \pm \frac{p^\mu}{p^2}
			= a^\mu \pm \frac{L}{2}\,\widehat{p}^\mu,
		\quad
			\widehat{p}^\mu = \frac{p^\mu}{\norm{p}}
		\end{equation}
		We can also nicely parametrize $X^I$ in terms of the proper length $\lambda$; for convenience, set $x^\mu(0) = a^\mu$, $z(0) = \frac{L}{2}$, then we have:
		\begin{equation}
			z(\lambda)
			= \frac{L}{2} \frac{1}{\cosh \lambda},
		\quad
			x^\mu(\lambda)
			= a^\mu \pm \widehat{p}^\mu \frac{L}{2}
				\tanh \lambda
		\label{eq:elliptic}
		\end{equation}

		\item For $p^2 < 0$, we have $(\frac{L}{2})^2 = \frac{1}{p^2} < 0$, and $\gamma$ is now a cross section of the two-sheet hyperboloid. Again as $z\to 0$ it lands at $x^\mu = a^\mu \pm \frac{L}{2}\,\widehat{p}^\mu$; however, $\abs{x^\mu}$ grows with $z$ and extends into the bulk instead of returning to the boundary, i.e.~$z,\abs{x^\mu} \to \infty$.

@@

	On the other hand, $\eta$ should be treated like an einbein: $\eta = \sqrt{-\gamma}$, here $\gamma = \gamma_{\tau\tau}$ is the worldline metric; $\gamma < 0$ due to the Lorentzian signature. We have:
	\begin{gather}
		\eta = \sqrt{-\gamma}
		\longmapsto \eta' = \eta\,
			\det \pdv{\tau}{\tau'}
		= \eta \pdv{\tau}{\tau'},
	\\[1ex]
		\eta^{-1}
		= - \sqrt{-\gamma}\,\gamma^{-1}
		\longmapsto (\eta')^{-1} = \eta^{-1}
			\pqty{\pdv{\tau}{\tau'}}
			\pqty{
				\pdv{\tau'}{\tau}
				\pdv{\tau'}{\tau}
			}
		= \eta^{-1} \pdv{\tau'}{\tau}
	\end{gather}
	It is clear that the action is invariant under the transformation:
	\begin{equation}
	\begin{aligned}
		S'
		&= \frac{1}{2} \int \dd{\tau'} \pqty\Big{
				(\eta')^{-1}
					\pdd{\tau'} X_\mu
					\pdd{\tau'} X^\mu
				- (\eta') m^2
			} \\
		&= \frac{1}{2} \int \dd{\tau}
			\pdv{\tau'}{\tau}\,
			\pqty{
				\eta^{-1}
					\pdv{\tau'}{\tau}
					\cdot
					\pdv{\tau}{\tau'}
					\pdd{\tau} X_\mu
					\cdot
					\pdv{\tau}{\tau'}
					\pdd{\tau} X^\mu
				- \eta\,\pdv{\tau}{\tau'}\,m^2
			} = S
	\end{aligned}
	\end{equation}

@@

\appendix

\section{Derivation of the special conformal transformations}
\label{sect:special_conformal}

>>>>

% !TeX document-id = {09950f11-703a-4175-9f6c-5ae00c774d6a}
% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX TXS-program:bibliography = biber -l zh__pinyin --output-safechars %

@@

\addbibresource{../gravity.bib}

@@

	From now on we shall focus on the $p^2 \ne 0$ situation. Note that we can complete the square on the right-hand side of the $k^\mu$ conservation such that:
	\begin{equation}
		k_\mu + p_\mu a^2
		= p_\mu \pqty{
			z^2 + (x - a)^2
		},
	\quad \Delta = p_\mu a^\mu
	\end{equation}
	For $p^\mu \ne 0$, we can always find some $c_\mu$ such that $c_\mu p^\mu \ne 0$, therefore we have:
	\begin{equation}
		z^2 + \pqty{x - a}^2
		= \frac{c_\mu k^\mu}{c_\mu p^\mu}
			+ a^2
		= \pqty{\frac{L}{2}}^{\!\!2}
		= \mrm{const.},
	\quad \Delta = p_\mu a^\mu
	\end{equation}

@@

	This confirms our observation above, and further reveals that $\gamma$ lies in a ``plane'' in the $\mbb{R}^d$ subspace consisting of the $x^\mu$ coordinates. The behavior of $\gamma$ is sensitive to the sign of $\delta$ and $p^2$; more specifically,
	\begin{enumerate}
	\item $\delta = -1$, i.e.~for timelike $\gamma$, we must have $p^2 < 0$, and $z \ge \frac{1}{\abs{\norm{p}}} = \frac{L}{2} > 0$, namely $z$ is bounded from below. This means that timelike geodesic can never reach the asymptotic boundary $z \to 0$. In this case, $\gamma$ is a section of the one-sheet hyperboloid. 
	\item $\delta = +1$, i.e.~for spacelike $\gamma$, we can have $p^2 > 0$ or $p^2 < 0$.
		\begin{enumerate}
		\item For $p^2 > 0$, again we have $\frac{L}{2} = \frac{1}{\norm{p}} > 0$, and $\gamma$ is again a cross-section of the one-sheet hyperboloid. However, now we have $z \le \frac{L}{2}$, namely $z$ is bounded from above, and:
		\begin{equation}
			z\to 0,
		\quad
			x^\mu
			= a^\mu \pm \frac{p^\mu}{p^2}
			= a^\mu \pm \frac{L}{2}\,\widehat{p}^\mu,
		\quad
			\widehat{p}^\mu = \frac{p^\mu}{\norm{p}}
		\end{equation}
		We can also nicely parametrize $X^I$ in terms of the proper length $\lambda$; for convenience, set $x^\mu(0) = a^\mu$, $z(0) = \frac{L}{2}$, then we have:
		\begin{equation}
			z(\lambda)
			= \frac{L}{2} \frac{1}{\cosh \lambda},
		\quad
			x^\mu(\lambda)
			= a^\mu \pm \widehat{p}^\mu \frac{L}{2}
				\tanh \lambda
		\label{eq:elliptic}
		\end{equation}

		\item For $p^2 < 0$, we have $(\frac{L}{2})^2 = \frac{1}{p^2} < 0$, and $\gamma$ is now a cross-section of the two-sheet hyperboloid. Again as $z\to 0$ it lands at $x^\mu = a^\mu \pm \frac{L}{2}\,\widehat{p}^\mu$; however, $\abs{x^\mu}$ grows with $z$ and extends into the bulk instead of returning to the boundary, i.e.~$z,\abs{x^\mu} \to \infty$.

@@

	On the other hand, $\eta$ should be treated like an einbein: $\eta = \sqrt{-\gamma}$, here $\gamma = \gamma_{\tau\tau}$ is the worldline metric; $\gamma < 0$ due to the Lorentzian signature \cite{Polchinski:1998rq}. We have:
	\begin{gather}
		\eta = \sqrt{-\gamma}
		\longmapsto \eta' = \eta\,
			\det \pdv{\tau}{\tau'}
		= \eta \pdv{\tau}{\tau'},
	\\[1ex]
		\eta^{-1}
		= - \sqrt{-\gamma}\,\gamma^{-1}
		\longmapsto (\eta')^{-1} = \eta^{-1}
			\pqty{\pdv{\tau}{\tau'}}
			\pqty{
				\pdv{\tau'}{\tau}
				\pdv{\tau'}{\tau}
			}
		= \eta^{-1} \pdv{\tau'}{\tau}
	\end{gather}
	It is clear that the action is invariant under the transformation:
	\begin{equation}
	\begin{aligned}
		S'
		&= \frac{1}{2} \int \dd{\tau'} \pqty\Big{
				(\eta')^{-1}
					\pdd{\tau'} X_\mu
					\pdd{\tau'} X^\mu
				- (\eta') m^2
			} \\
		&= \frac{1}{2} \int \dd{\tau}
			\pdv{\tau'}{\tau}\,
			\pqty{
				\eta^{-1}
					\pdv{\tau'}{\tau}
					\cdot
					\pdv{\tau}{\tau'}
					\pdd{\tau} X_\mu
					\cdot
					\pdv{\tau}{\tau'}
					\pdd{\tau} X^\mu
				- \eta\,\pdv{\tau}{\tau'}\,m^2
			} = S
	\end{aligned}
	\end{equation}

@@

	To find the connection form, we first note the \textit{Cartan's structure equation} \cite{Carroll:1997ar}:
	\begin{equation}
	\begin{aligned}
		\dd{e^a}
		= \dd{\pqty{e^a_\nu \dd{x^\nu}}}
		&= \pdd{\mu} e^a_\nu
			\dd{x^\mu} \wedge \dd{x^\nu} \\
		&= \pqty{
				- \omega^{a}_{\mu b} e^b_\nu
				+ \Gamma^{\lambda}_{\mu\nu} e^a_\lambda
			}
			\dd{x^\mu} \wedge \dd{x^\nu} \\
		&= -\omega\id{^a_b} \wedge e^b
			+ T^a
	\end{aligned}
	\end{equation}
	Where $T^a$ is the torsion tensor. 
	Here we've used the relation between spin connection $\omega^a_{\mu b}$ and the affine connection $\Gamma^\lambda_{\mu\nu}$. For the torsion-free Levi-Civita connection, we have $
		\dd{e^a}
		= -\omega\id{^a_b} \wedge e^b
	$.

	Rigidity of the veilbein, namely $
		\cdv{\mu} \pqty{g_{\mu\nu} e^\mu_a e^\nu_b}
		= \cdv{\mu} \eta_{ab} = 0
	$, further implies that:
	\begin{equation}
		\omega_{\mu ab}
		= -\omega_{\mu ba}
	\end{equation}
	With these constraints we can solve for the connection form, and we find that only the following components are non-vanishing\footnote{
		Go to \url{https://github.com/bryango/Archive/blob/master/HW-Gravity/gravity1/nb/vielbein.wl} for a Mathematica script for this calculation.
	}:
	\begin{gather}\allowdisplaybreaks
		\omega\id{^t_r}
		= \frac{f'}{2\sqrt{fh}} \dd{t}
		= -\omega\id{^r_t}
	\\
		\omega\id{^r_\theta}
		= \frac{-1}{\sqrt{h}} \dd{\theta}
		= -\omega\id{^\theta_r}
	\\
		\omega\id{^r_\phi}
		= \frac{-\sin\theta}{\sqrt{h}} \dd{\phi}
		= -\omega\id{^\phi_r}
	\\
		\omega\id{^\theta_\phi}
		= -\cos\theta \dd{\phi}
		= -\omega\id{^\phi_\theta}
	\end{gather}

	The curvature form is thus given by:
	\begin{equation}
		\Omega\id{^a_b}
		= \dd{\omega\id{^a_b}}
			+ \omega\id{^a_c} \wedge \omega\id{^c_b}
	\end{equation}
	The exterior derivative is easily computed in the $\dd{x^\mu}$ basis; however, to get to the Ricci tensor, it would be convenient to switch to the $e^a$ basis. Here we only write down the upper half of $\Omega\id{^a_b}$ since the lower half can be inferred from anti-symmetry:
	\begin{equation}
	\Omega\id{^a_b}
	\sim
	\pqty{
	\begin{array}{cccc}
		0
		& R\id{^t_r_t_r}\,e^t\wedge e^r
		& R\id{^t_\theta_t_\theta}\,e^t\wedge e^\theta
		& R\id{^t_\phi_t_\phi}\,e^t\wedge e^\phi \\
		\cdots & 0
		& R\id{^r_\theta_r_\theta}\,e^r\wedge e^\theta
		& R\id{^r_\phi_r_\phi}\,e^r\wedge e^\phi \\
		\cdots & \cdots & 0
		& R\id{^\theta_\phi_\theta_\phi}\,e^\theta\wedge e^\phi \\
		\cdots & \cdots & \cdots & 0 \\
	\end{array}
	}
	\end{equation}
	\begin{equation}
	\begin{aligned}
		R\id{^t_r_t_r}
		&= \frac{f'^2}{4f^2 h}
			+ \frac{f'h'}{4fh^2}
			- \frac{f''}{2fh},
	\\
		R\id{^t_\theta_t_\theta}
		&= \frac{-f'}{2rfh}
		= R\id{^t_\phi_t_\phi},
	\\
		R\id{^r_\theta_r_\theta}
		&= \frac{h'}{2rh^2}
		= R\id{^r_\phi_r_\phi},
	\\
		R\id{^\theta_\phi_\theta_\phi}
		&= \frac{h - 1}{r^2 h}
	\end{aligned}
	\end{equation}

	The Ricci tensor in $e^a$ basis is given by $
		R_{ab} = R\id{^c_a_c_b}
	$. Note that the components are non-zero iff.~$a = b$, i.e.~the Ricci tensor is diagonal. We have:
	\begin{equation}
	\begin{aligned}
		R_{ab}
		&\sim \mop{diag} \pqty\big{
			R\id{^r_t_r_t}
			+ R\id{^\theta_t_\theta_t}
			+ R\id{^\phi_t_\phi_t},
			\cdots,\cdots,\cdots
		} \\
		&= \mop{diag} \pqty\big{
			R\id{^t_r_t_r}
			+ R\id{^t_\theta_t_\theta}
			+ R\id{^t_\phi_t_\phi},
			\cdots,\cdots,\cdots
		} \\
		&= \mop{diag} \Big(\,
			R\id{^t_r_t_r}
			+ R\id{^t_\theta_t_\theta}
			+ R\id{^t_\phi_t_\phi},
		\\[-1ex] & \hspace{4.5em}
			R\id{^t_r_t_r}
			+ R\id{^r_\theta_r_\theta}
			+ R\id{^r_\phi_r_\phi},
		\\[-.5ex] & \hspace{4.5em}
			R\id{^t_\theta_t_\theta}
			+ R\id{^r_\theta_r_\theta}
			+ R\id{^\theta_\phi_\theta_\phi},
		\\[-1ex] & \hspace{4.5em}
			R\id{^t_\phi_t_\phi}
			+ R\id{^r_\phi_r_\phi}
			+ R\id{^\theta_\phi_\theta_\phi}
		\,\Big) \\
		&= \mop{diag} \Big(\,
			R\id{^t_r_t_r}
			+ 2R\id{^t_\theta_t_\theta},
		\\[-1ex] & \hspace{4.5em}
			R\id{^t_r_t_r}
			+ 2R\id{^r_\theta_r_\theta},
		\\[-.5ex] & \hspace{4.5em}
			R_{\theta\theta} = R\id{^t_\theta_t_\theta}
			+ R\id{^r_\theta_r_\theta}
			+ R\id{^\theta_\phi_\theta_\phi},
		\\[-1ex] & \hspace{4.5em}
			R_{\phi\phi} = R_{\theta\theta}
		\,\Big) \\
		&= \mop{diag} \bigg(\,
			\frac{f'^2}{4f^2 h}
			+ \frac{f'h'}{4fh^2}
			- \frac{f''}{2fh}
			- \frac{f'}{rfh},
		\\[-0ex] & \hspace{4.5em}
			\frac{f'^2}{4f^2 h}
			+ \frac{f'h'}{4fh^2}
			- \frac{f''}{2fh}
			+ \frac{h'}{rh^2},
		\\[-0ex] & \hspace{4.5em}
			R_{\theta\theta}
			= - \frac{f'}{2rfh}
			+ \frac{h'}{2rh^2}
			+ \frac{h - 1}{r^2 h},
		\\[-0ex] & \hspace{4.5em}
			R_{\phi\phi} = R_{\theta\theta}
		\,\bigg),
	\end{aligned}
	\end{equation}
	To go back to the $\dd{x^\mu}$ basis, we have $R_{\mu\nu} = R_{ab} e^a_\mu e^b_\nu$. In particular, here we have $R_{\mu\mu} = R_{aa} (e^a_\mu)^2$, therefore:
	\begin{equation}
	\begin{aligned}
		R_{\mu\nu}
		&= \mop{diag} \bigg(\,
			\frac{f'^2}{4f h}
			+ \frac{f'h'}{4h^2}
			- \frac{f''}{2h}
			- \frac{f'}{rh},
		\\[-0ex] & \hspace{4.5em}
			\frac{f'^2}{4f^2}
			+ \frac{f'h'}{4fh}
			- \frac{f''}{2f}
			+ \frac{h'}{rh},
		\\[-0ex] & \hspace{4.5em}
			R_{\theta\theta}
			= - \frac{rf'}{2fh}
			+ \frac{rh'}{2h^2}
			+ \frac{h - 1}{h},
		\\[-0ex] & \hspace{4.5em}
			R_{\phi\phi} = R_{\theta\theta} \sin^2\theta
		\,\bigg)
	\end{aligned}
	\end{equation}

\section{Jackiw--Teitelboim Gravity}
	The Jackiw--Teitelboim (JT) action is given by:
	\begin{equation}
		S = \frac{1}{16\pi G}
			\int \dd[2]{x} \sqrt{-g}\,\Phi\,
			\pqty{R + 2}
	\end{equation}
	$\fdv{S}{\Phi} = 0$ gives us $R = -2$. Now consider $\fdv{S}{g^{\mu\nu}}$, and we have:
	\begin{equation}
	\begin{aligned}
		\var_g {S}
		&= \frac{1}{16\pi G} \int \dd[2]{x}
			\Phi\,\var \pqty\Big{
				\sqrt{-g}\,\pqty{R + 2}
			},
		\quad R = g^{\mu\nu} R_{\mu\nu}, \\
		&= \frac{1}{16\pi G} \int \dd[2]{x} \sqrt{-g}\,
			\Phi\,\Bqty{
				\pqty{
					R_{\mu\nu} - \frac{1}{2} g_{\mu\nu} (R + 2)
				} \var{g^{\mu\nu}}
				+ g^{\mu\nu} \var{R_{\mu\nu}}
			}
	\end{aligned}
	\end{equation}

	Note that the $
		g^{\mu\nu} \var{R_{\mu\nu}}
	$ term is a total derivative\footnote{
		See the amazing lecture note by Matthias Blau at \url{http://www.blau.itp.unibe.ch/GRLecturenotes.html}.
	}:
	\begin{equation}
	\begin{aligned}
		g^{\mu\nu} \var{R_{\mu\nu}}
		&= \pqty{
				\nabla^\mu \nabla^\nu
				- g^{\mu\nu} \nabla^\lambda \nabla_\lambda
			} \var{g_{\mu\nu}} \\
		&= - \pqty{
				\cdv{\mu} \cdv{\nu}
				- g_{\mu\nu} \nabla^\lambda \nabla_\lambda
			} \var{g^{\mu\nu}} \\
		&= \cdv{\lambda} \pqty\Big{
				g^{\mu\nu} \var{\Gamma^\lambda_{\mu\nu}}
				- g^{\mu\lambda} \var{\Gamma^\nu_{\nu\mu}}
			}
	\end{aligned}
	\end{equation}
	In Einstein gravity this gets reduced to a boundary term. But here we have an additional factor of $\Phi$, so after integration by parts, we actually get the equation of motion (EoM) for $\Phi$, up to some boundary terms\footnote{
		See e.g.~Section 2 of \cite{Harlow:2018tqv}. 
	}:
	\begin{equation}
	\begin{aligned}
		\var_g {S}
		\sim \frac{1}{16\pi G} \int \dd[2]{x} \sqrt{-g}\,
			\Bqty{
				\Phi\,\pqty{
					R_{\mu\nu}
					- \frac{1}{2} g_{\mu\nu} (R + 2)
				} - \pqty\Big{
					\cdv{\mu} \cdv{\nu}
					- g_{\mu\nu} \nabla^\lambda \nabla_\lambda
				} \Phi
			} \var{g^{\mu\nu}}
	\end{aligned}
	\end{equation}

	In 2D, we have $
		R_{\mu\nu}
		\equiv \frac{1}{2} g_{\mu\nu} R
	$, so the EoM is simply:
	\begin{equation}
		\pqty{
			\cdv{\mu} \cdv{\nu}
			- g_{\mu\nu} \nabla^\lambda \nabla_\lambda
			+ g_{\mu\nu}
		}\, \Phi = 0
	\end{equation}
	Contraction with $g^{\mu\nu}$ further gives us $
		\pqty{
			\nabla^\lambda \nabla_\lambda - 2
		} \Phi = 0
	$, so in the end we have:
	\begin{equation}
		\pqty{
			\cdv{\mu} \cdv{\nu}
			- g_{\mu\nu}
		}\, \Phi = 0,
	\quad
		R = -2
	\end{equation}

\pagebreak
\appendix
\section{Derivation of the special conformal transformations}
\label{sect:special_conformal}
